---
title: "fig4"
output: html_document
date: "2025-04-12"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# 安装并加载 here 包
# install.packages("here")
library(here)

# Clean workspace
rm(list=ls()) 
# Load setting and functions
source("../Commonly_Used_R_Packages.R")
# Set output directory
output_dir="./"
```

## (A) PERMANOVA results of fecal metabolism profile
## Data source: fig4_A.csv
## Plot generated using GraphPad Prism 9

## (B) PCA of fecal metabolism profile
## Data source: fig4_B.csv
## Plot generated using R

```{r}

# Load necessary library
library(dplyr)

# Read and preprocess the data
pca_coordinate  <- read.csv("fig4_B.csv", header = 1, check.names = F) %>%
  mutate(
    group = as.factor(substr(subject, 1, 1)),
    day_number = as.numeric(gsub("D", "", Day0_27)),
    period = case_when(
      day_number >= 0 & day_number <= 13 ~ "Normal",
      day_number >= 14 & day_number <= 20 ~ "Low-dose",
      day_number >= 21 & day_number <= 27 ~ "High-dose",
      day_number == 56 ~ "Follow-up",
      TRUE ~ NA_character_ 
    )
  )

p_period =ggplot(pca_coordinate,aes(x=PC3,y=PC4,color=period, shape= group))+   
  geom_point(size=2.2) +
  scale_color_manual(name = "Period",
                     values = c("#737377","#CBE04C","#6EBC10", "#BFBFC1"))+
  guides(colour = guide_legend(ncol = 1, override.aes = list(size=5)))+
  scale_shape_manual(name = "Group",values = c(16, 15)) +
  theme_bw() +
  theme(panel.border=element_blank(),
        panel.grid.major=element_blank(),
        panel.grid.minor=element_blank(),
        axis.line= element_line(colour = "black"),
        axis.text.y = element_text(size = 16,colour="black"),#face="bold",bold表示加粗
        axis.title = element_text(size = 16,colour="black"),
        legend.text=element_text(size = 16),
        legend.key=element_blank(),
        legend.position = "None",
        plot.title = element_text(size = 16),
        axis.text.x=element_text(size = 16,colour="black"))+
  xlab("PC3(4.07%)") +
  ylab("PC4(3.73%)")+
  theme(text=element_text(size=16))
p_period

```

## (C) Boxplot of PC3 values across different phases
## Data source: fig4_C.csv
## Plot generated using GraphPad Prism 9

## (D) Boxplot of PC4 values across different phases
## Data source: fig4_D.csv
## Plot generated using GraphPad Prism 9

## (E) Line plot of daily PC3 values throughout the entire trial
## Data source: fig4_E.csv
## Plot generated using GraphPad Prism 9

## (F) Line plot of daily PC4 values throughout the entire trial
## Data source: fig4_F.csv
## Plot generated using GraphPad Prism 9

## (G) Upregulated metabolic pathways during low-dose fiber intervention (vs. normal diet) in each subject of the Overweight group
## Data source: fig4_G.csv
## Plot generated using R

```{r}

# Clean workspace
rm(list=ls()) 

print_Df  <- read.csv("fig4_G.csv")  %>%
  mutate(
    ID = as.factor(ID),
    pathway = as.factor(pathway),
    
    enrichment_radio = as.numeric(enrichment_radio.x),
    Raw_p = as.numeric(Raw_p)
  )

p <- ggplot(print_Df, aes(ID, pathway)) +
  geom_point(aes(color = -log10(Raw_p), size = enrichment_radio)) +
  scale_size_continuous(name = "Enrichment radio",
                        range = c(0, 17.586),  # 设置点的大小范围
                        limits = c(1, 100))+
  scale_color_gradientn(name = "-log10(p value)", 
                        colors = c('#55B047', '#FBA304', '#FF1900'),
                        limits = c(0, 4.2),
                        breaks = c(0, 1, 2, 3, 4), 
                        labels = c("0", "1", "2", "3", "4")) +
  theme_bw() +
  labs(x = '', y = '')+
  theme(
    axis.line= element_line(colour = "black"),
    axis.text.y = element_text(size = 16,colour="black"),#face="bold",bold表示加粗
    axis.title = element_text(size = 16,colour="black"),
    axis.text.x = element_text(size = 16,colour="black"),
    legend.text=element_text(size = 16),
    legend.key=element_blank(),
    legend.position = "none",
    plot.title = element_text(size = 16))+
  theme(text=element_text(size=16,  family="sans"))
p

```

## (H) Upregulated metabolic pathways during high-dose fiber intervention (vs. low-dose fiber) in each subject of the Overweight group
## Data source: fig4_H.csv
## Plot generated using R

```{r, echo=FALSE, fig.width=20, fig.height=3}

# Clean workspace
rm(list=ls()) 

# Load and preprocess data
print_Df  <- read.csv("fig4_H.csv")  %>%
  mutate(
    ID = as.factor(ID),
    pathway = as.factor(pathway),
    
    enrichment_radio = as.numeric(enrichment_radio.x),
    Raw_p = as.numeric(Raw_p)
  )

# Compute pathway order: by count of valid p-values and average p-value
pathway_order <- print_Df %>%
  group_by(pathway) %>%
  summarise(
    n = sum(!is.na(Raw_p)),    # Number of valid p-values
    avg_p = mean(Raw_p, na.rm = TRUE)  # Mean p-value 
  ) %>%
  arrange(desc(n), avg_p) # Sort: more frequent first, then lower p-value

# Set pathway as ordered factor (most significant on top)
print_Df <- print_Df %>%
  mutate(pathway = factor(pathway, levels = rev(pathway_order$pathway)))

# Plot dotplot
p <- ggplot(print_Df, aes(ID, pathway)) +
  geom_point(aes(color = -log10(Raw_p), size = enrichment_radio)) +
  scale_size_continuous(name = "Enrichment radio",
                        range = c(0, 17.586),  # 设置点的大小范围
                        limits = c(1, 100))+
  scale_color_gradientn(name = "-log10(p value)", 
                        colors = c('#55B047', '#FBA304', '#FF1900'),
                        limits = c(0, 4.2),
                        breaks = c(0, 1, 2, 3, 4), 
                        labels = c("0", "1", "2", "3", "4")) +
  theme_bw() +
  labs(x = '', y = '')+
  theme(
    axis.line= element_line(colour = "black"),
    axis.text.y = element_text(size = 16,colour="black"),#face="bold",bold表示加粗
    axis.title = element_text(size = 16,colour="black"),
    axis.text.x = element_text(size = 16,colour="black"),
    legend.text=element_text(size = 16),
    legend.key=element_blank(),
    legend.position = "none",
    plot.title = element_text(size = 16))+
  theme(text=element_text(size=16,  family="sans"))
p

```

## (I) Upregulated metabolic pathways during high-dose fiber intervention (vs. normal diet) in each subject of the Overweight group
## Data source: fig4_I.csv
## Plot generated using R

```{r, echo=FALSE, fig.width=20, fig.height=4}

# Clean workspace
rm(list=ls()) 

# Load and preprocess data
print_Df  <- read.csv("fig4_I.csv")  %>%
  mutate(
    ID = as.factor(ID),
    pathway = as.factor(pathway),
    
    enrichment_radio = as.numeric(enrichment_radio.x),
    Raw_p = as.numeric(Raw_p)
  )

# Compute pathway order: by count of valid p-values and average p-value
pathway_order <- print_Df %>%
  group_by(pathway) %>%
  summarise(
    n = sum(!is.na(Raw_p)),    # Number of valid p-values
    avg_p = mean(Raw_p, na.rm = TRUE)  # Mean p-value 
  ) %>%
  arrange(desc(n), avg_p) # Sort: more frequent first, then lower p-value

# Set pathway as ordered factor (most significant on top)
print_Df <- print_Df %>%
  mutate(pathway = factor(pathway, levels = rev(pathway_order$pathway)))

# Plot dotplot
p <- ggplot(print_Df, aes(ID, pathway)) +
  geom_point(aes(color = -log10(Raw_p), size = enrichment_radio)) +
  scale_size_continuous(name = "Enrichment radio",
                        range = c(0, 17.586),  # 设置点的大小范围
                        limits = c(1, 100))+
  scale_color_gradientn(name = "-log10(p value)", 
                        colors = c('#55B047', '#FBA304', '#FF1900'),
                        limits = c(0, 4.2),
                        breaks = c(0, 1, 2, 3, 4), 
                        labels = c("0", "1", "2", "3", "4")) +
  theme_bw() +
  labs(x = '', y = '')+
  theme(
    axis.line= element_line(colour = "black"),
    axis.text.y = element_text(size = 16,colour="black"),#face="bold",bold表示加粗
    axis.title = element_text(size = 16,colour="black"),
    axis.text.x = element_text(size = 16,colour="black"),
    legend.text=element_text(size = 16),
    legend.key=element_blank(),
    legend.position = "none",
    plot.title = element_text(size = 16))+
  theme(text=element_text(size=16,  family="sans"))
p

```

## (J) Upregulated metabolic pathways during low-dose fiber intervention (vs. normal diet) in each subject of the T2D group
## Data source: fig4_J.csv
## Plot generated using R

```{r, echo=FALSE, fig.width=10, fig.height=2}

# Clean workspace
rm(list=ls()) 

# Load and preprocess data
print_Df  <- read.csv("fig4_J.csv")  %>%
  mutate(
    ID = as.factor(ID),
    pathway = as.factor(pathway),
    
    enrichment_radio = as.numeric(enrichment_radio.x),
    Raw_p = as.numeric(Raw_p)
  )

# Compute pathway order: by count of valid p-values and average p-value
pathway_order <- print_Df %>%
  group_by(pathway) %>%
  summarise(
    n = sum(!is.na(Raw_p)),    # Number of valid p-values
    avg_p = mean(Raw_p, na.rm = TRUE)  # Mean p-value 
  ) %>%
  arrange(desc(n), avg_p) # Sort: more frequent first, then lower p-value

# Set pathway as ordered factor (most significant on top)
print_Df <- print_Df %>%
  mutate(pathway = factor(pathway, levels = rev(pathway_order$pathway)))

# Plot dotplot
p <- ggplot(print_Df, aes(ID, pathway)) +
  geom_point(aes(color = -log10(Raw_p), size = enrichment_radio)) +
  scale_size_continuous(name = "Enrichment radio",
                        range = c(0, 17.586),  # 设置点的大小范围
                        limits = c(1, 100))+
  scale_color_gradientn(name = "-log10(p value)", 
                        colors = c('#55B047', '#FBA304', '#FF1900'),
                        limits = c(0, 4.2),
                        breaks = c(0, 1, 2, 3, 4), 
                        labels = c("0", "1", "2", "3", "4")) +
  theme_bw() +
  labs(x = '', y = '')+
  theme(
    axis.line= element_line(colour = "black"),
    axis.text.y = element_text(size = 16,colour="black"),#face="bold",bold表示加粗
    axis.title = element_text(size = 16,colour="black"),
    axis.text.x = element_text(size = 16,colour="black"),
    legend.text=element_text(size = 16),
    legend.key=element_blank(),
    legend.position = "none",
    plot.title = element_text(size = 16))+
  theme(text=element_text(size=16,  family="sans"))
p

```

## (K) Upregulated metabolic pathways during high-dose fiber intervention (vs. low-dose fiber) in each subject of the T2D group
## Data source: fig4_K.csv
## Plot generated using R

```{r, echo=FALSE, fig.width=15, fig.height=3}

# Clean workspace
rm(list=ls()) 

# Load and preprocess data
print_Df  <- read.csv("fig4_K.csv")  %>%
  mutate(
    ID = as.factor(ID),
    pathway = as.factor(pathway),
    
    enrichment_radio = as.numeric(enrichment_radio.x),
    Raw_p = as.numeric(Raw_p)
  )

# Compute pathway order: by count of valid p-values and average p-value
pathway_order <- print_Df %>%
  group_by(pathway) %>%
  summarise(
    n = sum(!is.na(Raw_p)),    # Number of valid p-values
    avg_p = mean(Raw_p, na.rm = TRUE)  # Mean p-value 
  ) %>%
  arrange(desc(n), avg_p) # Sort: more frequent first, then lower p-value

# Set pathway as ordered factor (most significant on top)
print_Df <- print_Df %>%
  mutate(pathway = factor(pathway, levels = rev(pathway_order$pathway)))

# Plot dotplot
p <- ggplot(print_Df, aes(ID, pathway)) +
  geom_point(aes(color = -log10(Raw_p), size = enrichment_radio)) +
  scale_size_continuous(name = "Enrichment radio",
                        range = c(0, 17.586),  # 设置点的大小范围
                        limits = c(1, 100))+
  scale_color_gradientn(name = "-log10(p value)", 
                        colors = c('#55B047', '#FBA304', '#FF1900'),
                        limits = c(0, 4.2),
                        breaks = c(0, 1, 2, 3, 4), 
                        labels = c("0", "1", "2", "3", "4")) +
  theme_bw() +
  labs(x = '', y = '')+
  theme(
    axis.line= element_line(colour = "black"),
    axis.text.y = element_text(size = 16,colour="black"),#face="bold",bold表示加粗
    axis.title = element_text(size = 16,colour="black"),
    axis.text.x = element_text(size = 16,colour="black"),
    legend.text=element_text(size = 16),
    legend.key=element_blank(),
    legend.position = "none",
    plot.title = element_text(size = 16))+
  theme(text=element_text(size=16,  family="sans"))
p

```

## (L) Upregulated metabolic pathways during high-dose fiber intervention (vs. normal diet) in each subject of the T2D group
## Data source: fig4_L.csv
## Plot generated using R

```{r, echo=FALSE, fig.width=15, fig.height=5}

# Clean workspace
rm(list=ls()) 

# Load and preprocess data
print_Df  <- read.csv("fig4_L.csv")  %>%
  mutate(
    ID = as.factor(ID),
    pathway = as.factor(pathway),
    
    enrichment_radio = as.numeric(enrichment_radio.x),
    Raw_p = as.numeric(Raw_p)
  )

# Compute pathway order: by count of valid p-values and average p-value
pathway_order <- print_Df %>%
  group_by(pathway) %>%
  summarise(
    n = sum(!is.na(Raw_p)),    # Number of valid p-values
    avg_p = mean(Raw_p, na.rm = TRUE)  # Mean p-value 
  ) %>%
  arrange(desc(n), avg_p) # Sort: more frequent first, then lower p-value

# Set pathway as ordered factor (most significant on top)
print_Df <- print_Df %>%
  mutate(pathway = factor(pathway, levels = rev(pathway_order$pathway)))

# Plot dotplot
p <- ggplot(print_Df, aes(ID, pathway)) +
  geom_point(aes(color = -log10(Raw_p), size = enrichment_radio)) +
  scale_size_continuous(name = "Enrichment radio",
                        range = c(0, 17.586),  # 设置点的大小范围
                        limits = c(1, 100))+
  scale_color_gradientn(name = "-log10(p value)", 
                        colors = c('#55B047', '#FBA304', '#FF1900'),
                        limits = c(0, 4.2),
                        breaks = c(0, 1, 2, 3, 4), 
                        labels = c("0", "1", "2", "3", "4")) +
  theme_bw() +
  labs(x = '', y = '')+
  theme(
    axis.line= element_line(colour = "black"),
    axis.text.y = element_text(size = 16,colour="black"),#face="bold",bold表示加粗
    axis.title = element_text(size = 16,colour="black"),
    axis.text.x = element_text(size = 16,colour="black"),
    legend.text=element_text(size = 16),
    legend.key=element_blank(),
    legend.position = "none",
    plot.title = element_text(size = 16))+
  theme(text=element_text(size=16,  family="sans"))
p

```