---
title: "fig5"
output: html_document
date: "2025-04-12"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Clean workspace
rm(list=ls()) 
# Load setting and functions
source("../Commonly_Used_R_Packages.R")
# Set output directory
output_dir="./"
```

## (A) Time-delay network between fecal metabolite concentrations and CAG abundances
## Data source: fig5_A.csv
## Plot generated using Cytoscape

## (B) Daily abundance changes of CAG80 during the intervention period
## Data source: fig5_B.csv
## Plot generated using R

```{r, echo=FALSE, fig.width=10, fig.height=4}

group_color <-c("#FDD0A2", "#FDBD83", "#FDAA65", "#FD984B", "#FA8532", "#F3711C", "#E95E0D", "#DB4B03", "#C24001", "#A63603")

df <-  read.csv("fig5_B.csv",header = 1, check.names = F) %>%
  mutate(
    group = as.factor(substr(ID, 1, 1)),
    day_number = as.numeric(gsub("D", "", Day0_27)),
    period = case_when(
      day_number >= 0 & day_number <= 13 ~ "Normal",
      day_number >= 14 & day_number <= 20 ~ "Low-dose",
      day_number >= 21 & day_number <= 27 ~ "High-dose",
      day_number == 56 ~ "Follow-up",
      TRUE ~ NA_character_ 
    )
  )
# head(df)
## 绘图
library(ggplot2)

p <- ggplot(df, aes(
  x = day_number,
  y = `O_CAG80_perc`,
  color = as.character(ID),
  group = ID
)) +
  geom_point(size = 1.5, na.rm = TRUE) +
  geom_line(na.rm = TRUE) +
  scale_x_continuous(
    expand = c(0, 0),
    breaks = 0:28,
    limits = c(0, 28)
  ) +
  scale_color_manual("Subject", values = group_color) +
  labs(
    x = "Day",
    y = "Relative conc of\nO_CAG80"
  ) +
  guides(colour = guide_legend(ncol = 6)) +
  theme(
    text = element_text(size = 14, family = "sans"),  # 全局字体设置
    axis.title = element_text(size = 12, colour = "black"),
    axis.text = element_text(size = 12, colour = "black"),
    axis.line = element_line(colour = "black"),
    panel.grid = element_blank(),
    panel.background = element_blank(),
    legend.text = element_text(size = 12),
    legend.key = element_blank(),
    plot.title = element_text(size = 12),
    legend.position = "top"
  )

p
```

## (C) Daily concentration changes of 4-Pyridoxic acid during the intervention period
## Data source: fig5_C.csv
## Plot generated using R

```{r, echo=FALSE, fig.width=10, fig.height=4}
group_color <-c("#FDD0A2", "#FDBD83", "#FDAA65", "#FD984B", "#FA8532", "#F3711C", "#E95E0D", "#DB4B03", "#C24001", "#A63603")

df <-  read.csv("fig5_C.csv",header = 1, check.names = F) %>%
  mutate(
    group = as.factor(substr(ID, 1, 1)),
    day_number = as.numeric(gsub("D", "", Day0_27)),
    period = case_when(
      day_number >= 0 & day_number <= 13 ~ "Normal",
      day_number >= 14 & day_number <= 20 ~ "Low-dose",
      day_number >= 21 & day_number <= 27 ~ "High-dose",
      day_number == 56 ~ "Follow-up",
      TRUE ~ NA_character_ 
    )
  )
## 绘图
library(ggplot2)

## 绘图
library(ggplot2)

p <- ggplot(df, aes(
  x = day_number,
  y = `4-Pyridoxic_acid_perc`,
  color = as.character(ID),
  group = ID
)) +
  geom_point(size = 1.5, na.rm = TRUE) +
  geom_line(na.rm = TRUE) +
  scale_x_continuous(
    expand = c(0, 0),
    breaks = 0:28,
    limits = c(0, 28)
  ) +
  scale_color_manual("Subject", values = group_color) +
  labs(
    x = "Day",
    y = "Relative conc of\n4-Pyridoxic acid"
  ) +
  guides(colour = guide_legend(ncol = 6)) +
  theme(
    text = element_text(size = 14, family = "sans"),  # 全局字体设置
    axis.title = element_text(size = 12, colour = "black"),
    axis.text = element_text(size = 12, colour = "black"),
    axis.line = element_line(colour = "black"),
    panel.grid = element_blank(),
    panel.background = element_blank(),
    legend.text = element_text(size = 12),
    legend.key = element_blank(),
    plot.title = element_text(size = 12),
    legend.position = "top"
  )

p
```

## (D) Metabolic flow chart: Vitamin B6 metabolism (KEGG pathway #00750)
## Data source: fig5_D.csv
## Plot generated using Adobe Illustrator

## (E) Network of relationships among HQMAGs within CAG80
## Data source: fig5_E.csv
## Plot generated using Cytoscape

## (F) Repeated-measures correlation between 4-Pyridoxic acid and CGM metrics
## Data source: fig5_F.csv
## Plot generated using R

```{r, echo=FALSE, fig.width=5, fig.height=6}
# Clean workspace
rm(list = ls()) 

# Set group label and color scale range
group_print <- "Overweight"
bk_set <- c(-0.2, 0, 0.2)

# Load correlation and p-value matrix
rmcorr_df <- read.csv("fig5_F.csv", header = TRUE, check.names = FALSE, row.names = 1)

# Convert p-values to significance symbols
p_value <- rmcorr_df %>%
  select("P_value") %>%
  replace(. <= 0.001, "***") %>%
  replace(. >= 0.001 & . <= 0.01, "**") %>%
  replace(. >= 0.01 & . <= 0.05, "*") %>%
  replace(. > 0.05 & . <= 0.1, "+") %>%
  replace(. > 0.1, "")

# Extract correlation values only
rmcorr_df <- rmcorr_df %>%
  select("R_value")

# Load ComplexHeatmap for plotting
library(ComplexHeatmap)

# Create annotated heatmap
heatmap_object <- Heatmap(
  as.matrix(rmcorr_df),
  cluster_columns = FALSE,
  cluster_rows = FALSE,
  column_split = rep(group_print, times = ncol(rmcorr_df)),
  border_gp = gpar(col = "black", lwd = 0.3),
  rect_gp = gpar(col = "black", lwd = 0.3),
  column_labels = colnames(rmcorr_df),
  row_labels = rownames(rmcorr_df),
  column_names_rot = 90,
  column_names_gp = gpar(fontsize = 10),
  row_names_gp = gpar(fontsize = 10),
  row_dend_width = unit(2, "cm"),
  column_dend_height = unit(2, "cm"),
  show_column_dend = FALSE,
  cell_fun = function(j, i, x, y, width, height, fill) {
    grid.text(p_value[i, j], x, y, gp = gpar(col = "white", fontsize = 12))
  },
  width = unit(0.5, "cm"),
  height = unit(8, "cm"),
  col = colorRamp2(bk_set, c("#3174A1", "white", "#BF3D3D")),
  heatmap_legend_param = list(
    title = "Correlation",
    title_position = "topcenter",
    title_gp = gpar(fontsize = 10),
    at = bk_set,
    legend_direction = "vertical",
    grid_width = unit(2, "mm"),
    grid_height = unit(4, "mm"),
    labels_gp = gpar(fontsize = 10)
  )
)

# Render the heatmap
heatmap_object <- draw(heatmap_object)
```
