---
title: "fig2"
output: html_document
date: "2025-04-12"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# 安装并加载 here 包
# install.packages("here")
library(here)

# Clean workspace
rm(list=ls()) 
# Load setting and functions
source("../Commonly_Used_R_Packages.R")
# Set output directory
output_dir="./"
```

## (H) APCoA plot 
## Data source: fig2_H.csv
## Plot generated using R

```{r load-data, message=FALSE, warning=FALSE}
# Load necessary library
library(dplyr)

# Read and preprocess the data
aPCoA.score <- read.csv(here("fig2_H.csv")) %>%
  mutate(
    group = as.factor(substr(ID, 1, 1)),
    day_number = as.numeric(gsub("D", "", Day0_27)),
    period = case_when(
      day_number >= 0 & day_number <= 13 ~ "Normal",
      day_number >= 14 & day_number <= 20 ~ "Low-dose",
      day_number >= 21 & day_number <= 27 ~ "High-dose",
      day_number == 56 ~ "Follow-up",
      TRUE ~ NA_character_ 
    )
  )
```

```{r load-data, message=FALSE, warning=FALSE}
# Load ggplot2
library(ggplot2)

# Generate the PCoA plot
shape_NT <- ggplot(aPCoA.score, aes(PCoA1, PCoA2, color = period, shape = group, group = as.factor(ID))) +
  geom_point(size = 2.2) +
  scale_shape_manual(name = "Group", values = c(16, 15)) +
  scale_color_manual(name = "Period",
                     values = c("#737377", "#CBE04C", "#6EBC10", "#BFBFC1")) +
  theme_bw() +
  theme(
    panel.border = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.line = element_line(colour = "black"),
    axis.text = element_text(size = 16, colour = "black"),
    axis.title = element_text(size = 16, colour = "black"),
    legend.text = element_text(size = 16),
    legend.key = element_blank(),
    legend.position = "none",
    plot.title = element_text(size = 16),
    text = element_text(size = 16, family = "sans")
  ) +
  labs(
    x = paste('PC1:', '11.01', '%'),
    y = paste('PC2:', '7.41', '%')
  )

# Display the plot
shape_NT
```

## (A) Boxplot of Shannon index across different phages
## Data source: fig2_A.csv
## Plot generated using GraphPad Prism 9

## (B) Boxplot of strain number across different phages
## Data source: fig2_B.csv
## Plot generated using GraphPad Prism 9

## (C) Line plot of daily Shannon index throughout the entire trial
## Data source: fig2_C.csv
## Plot generated using GraphPad Prism 9

## (D) Line plot of daily strain number throughout the entire trial
## Data source: fig2_D.csv
## Plot generated using GraphPad Prism 9

## (E) Boxplot of coefficient of variation (CV) of Shannon index across different phages
## Data source: fig2_E.csv
## Plot generated using GraphPad Prism 9

## (F) Boxplot of coefficient of variation (CV) of strain number across different phages
## Data source: fig2_F.csv
## Plot generated using GraphPad Prism 9

## (G) PERMANOVA results comparing microbial communities
## Data source: fig2_G.csv
## Plot generated using GraphPad Prism 9

## (I) Boxplot of PC1 values across different phages
## Data source: fig2_I.csv
## Plot generated using GraphPad Prism 9

## (J) Boxplot of PC2 values across different phages
## Data source: fig2_J.csv
## Plot generated using GraphPad Prism 9

## (K) Line plot of daily PC1 values throughout the entire trial
## Data source: fig2_K.csv
## Plot generated using GraphPad Prism 9

## (L) Line plot of daily PC2 values throughout the entire trial
## Data source: fig2_L.csv
## Plot generated using GraphPad Prism 9