---
title: "fig3"
output: html_document
date: "2025-04-12"
knit: (function(inputFile, encoding) {
  knitr::opts_knit$set(root.dir = dirname(inputFile))
  rmarkdown::render(inputFile, encoding = encoding)
})
---

```{r setup}
knitr::opts_chunk$set(echo = TRUE)

# 安装并加载 here 包
# install.packages("here")
library(here)
library(dplyr)
library(ComplexHeatmap)
library(circlize)

# Clean workspace
rm(list=ls()) 

# Load setting and functions
source("../Commonly_Used_R_Packages.R")
# Set output directory
output_dir="./"
```

## (C) Daily abundance changes of CAGs during the intervention period in the Overweight group
## Data source: fig3_C.csv
## Plot generated using R and Adobe Illustrator

```{r}
# Clean workspace
rm(list=ls()) 

# Group name to be displayed
group_Set_print <- "Overwight"

# Load z-score normalized data and grouping info
sorted_group <- read.csv("fig3_C_z_score.csv", row.names = 1)
sig_stand_df <- sorted_group %>% select(-"Group")

# Load LMM p-values and annotate significance
signal_row <- read.csv("fig3_C_LMM_p_value.csv", row.names = 1, check.names = FALSE) %>%
  mutate_all(~ case_when(
    . < 0.001 ~ "***",
    . < 0.01  ~ "**",
    . < 0.05  ~ "*",
    . < 0.1   ~ "+",
    TRUE      ~ ""
  )) %>%
  mutate(Nor_mean = "") %>%
  select(Nor_mean, everything())

# Load Linda results and classify regulation direction
linda_df <- read.csv("fig3_C_Linda_p_value.csv")
full_linda_df <- data.frame(
  CAG = rownames(sig_stand_df),
  log2FoldChange = NA_real_,
  pvalue = NA_real_,
  stringsAsFactors = FALSE
) %>%
  rows_update(linda_df, by = "CAG") %>%
  mutate(label = case_when(
    pvalue <= 0.05 & log2FoldChange > 0 ~ "Up",
    pvalue <= 0.05 & log2FoldChange < 0 ~ "Down",
    TRUE ~ "NS"
  ))

# Define annotation colors
label_colors <- c(
  "Up" = "#BF3D3D",      # red
  "Down" = "#3174A1",    # blue
  "NS" = "white"         # not significant
)

# Build row annotation
ha <- HeatmapAnnotation(
  bar = full_linda_df$label,
  col = list(bar = label_colors),
  which = "row",
  simple_anno_size = unit(0.3, "cm")
)

# Set heatmap labels
column_split = rep(group_Set_print, times = ncol(sig_stand_df))
column_labels = rep(gsub("day", " ", colnames(sig_stand_df)[1:15]), times = 1)
row_labels = substring(rownames(sig_stand_df), 3)

# Create the heatmap object
heatmap_object <- Heatmap(as.matrix(sig_stand_df),
                          cluster_columns = FALSE,
                          cluster_rows = FALSE,
                          column_split = column_split,
                          border_gp = gpar(col = "black", lwd = 0.3),
                          rect_gp = gpar(col = "black", lwd = 0.3),
                          column_labels = column_labels,
                          row_labels = row_labels,
                          column_names_rot = 90,
                          column_names_gp = gpar(fontsize = 6),
                          row_names_gp = gpar(fontsize = 6),
                          row_dend_width = unit(2, "cm"),
                          column_dend_height = unit(2, "cm"),
                          show_column_dend = TRUE,
                          row_split = factor(sorted_group$Group),
                          row_title = "Cluster %s",
                          width = unit(5, "cm"),
                          height = unit(20, "cm"),
                          cell_fun = function(j, i, x, y, width, height, fill) {
                            grid.text(signal_row[i, j], x, y, gp = gpar(col = "white", fontsize = 6))
                          },
                          right_annotation = ha,
                          col = colorRamp2(c(-4.5, 0, 4.5), c("black", "white", "#DC8418")),
                          heatmap_legend_param = list(
                            title = "Correlation",
                            title_position = "topcenter",
                            title_gp = gpar(fontsize = 6, fontface = "plain"),
                            at = c(-4.5, 0, 4.5),
                            legend_position = 'top',
                            legend_direction = 'vertical',
                            grid_width = unit(2, "mm"),
                            grid_height = unit(4, "mm"),
                            labels_gp = gpar(fontsize = 6)
                          ))

# Draw the heatmap
draw(heatmap_object)
```

## (D) Daily abundance changes of CAGs during the intervention period in the T2D group
## Data source: fig3_D.csv
## Plot generated using R and Adobe Illustrator

```{r}
# Clean workspace
rm(list=ls()) 

# Group name to be displayed
group_Set_print <- "T2D"

# Load z-score normalized data and grouping info
sorted_group <- read.csv("fig3_D_z_score.csv", row.names = 1)
sig_stand_df <- sorted_group %>% select(-"Group")

# Load LMM p-values and annotate significance
signal_row <- read.csv("fig3_D_LMM_p_value.csv", row.names = 1, check.names = FALSE) %>%
  mutate_all(~ case_when(
    . < 0.001 ~ "***",
    . < 0.01  ~ "**",
    . < 0.05  ~ "*",
    . < 0.1   ~ "+",
    TRUE      ~ ""
  )) %>%
  mutate(Nor_mean = "") %>%
  select(Nor_mean, everything())

# Load Linda results and classify regulation direction
linda_df <- read.csv("fig3_D_Linda_p_value.csv")
full_linda_df <- data.frame(
  CAG = rownames(sig_stand_df),
  log2FoldChange = NA_real_,
  pvalue = NA_real_,
  stringsAsFactors = FALSE
) %>%
  rows_update(linda_df, by = "CAG") %>%
  mutate(label = case_when(
    pvalue <= 0.05 & log2FoldChange > 0 ~ "Up",
    pvalue <= 0.05 & log2FoldChange < 0 ~ "Down",
    TRUE ~ "NS"
  ))

# Define annotation colors
label_colors <- c(
  "Up" = "#BF3D3D",      # red
  "Down" = "#3174A1",    # blue
  "NS" = "white"         # not significant
)

# Build row annotation
ha <- HeatmapAnnotation(
  bar = full_linda_df$label,
  col = list(bar = label_colors),
  which = "row",
  simple_anno_size = unit(0.3, "cm")
)

# Set heatmap labels
column_split = rep(group_Set_print, times = ncol(sig_stand_df))
column_labels = rep(gsub("day", " ", colnames(sig_stand_df)[1:15]), times = 1)
row_labels = substring(rownames(sig_stand_df), 3)

# Create the heatmap object
heatmap_object <- Heatmap(as.matrix(sig_stand_df),
                          cluster_columns = FALSE,
                          cluster_rows = FALSE,
                          column_split = column_split,
                          border_gp = gpar(col = "black", lwd = 0.3),
                          rect_gp = gpar(col = "black", lwd = 0.3),
                          column_labels = column_labels,
                          row_labels = row_labels,
                          column_names_rot = 90,
                          column_names_gp = gpar(fontsize = 6),
                          row_names_gp = gpar(fontsize = 6),
                          row_dend_width = unit(2, "cm"),
                          column_dend_height = unit(2, "cm"),
                          show_column_dend = TRUE,
                          row_split = factor(sorted_group$Group),
                          row_title = "Cluster %s",
                          width = unit(5, "cm"),
                          height = unit(20, "cm"),
                          cell_fun = function(j, i, x, y, width, height, fill) {
                            grid.text(signal_row[i, j], x, y, gp = gpar(col = "white", fontsize = 6))
                          },
                          right_annotation = ha,
                          col = colorRamp2(c(-4.5, 0, 4.5), c("black", "white", "#DC8418")),
                          heatmap_legend_param = list(
                            title = "Correlation",
                            title_position = "topcenter",
                            title_gp = gpar(fontsize = 6, fontface = "plain"),
                            at = c(-4.5, 0, 4.5),
                            legend_position = 'top',
                            legend_direction = 'vertical',
                            grid_width = unit(2, "mm"),
                            grid_height = unit(4, "mm"),
                            labels_gp = gpar(fontsize = 6)
                          ))

# Draw the heatmap
draw(heatmap_object)
```

## (A) Co-abundance network for the Overweight group
## Data source: fig3_A.csv
## Plot generated using Cytoscape

## (B) Co-abundance network for the T2D group
## Data source: fig3_B.csv
## Plot generated using Cytoscape

## (E) CAZy gene characteristics of clusters with different abundance trends in the Overweight group
## Data source: fig3_E.csv
## Plot generated using R


```{r}
# Clean workspace
rm(list = ls()) 

# Set display group label and color settings
group_Set_print <- "Overweight"
colors <- c("black", "white", "#DC8418")
break_set <- c(-2.5, 0, 2.5)

# Load data
cluster_z_score <- read.csv("fig3_E_GeneProportion.csv")
mw_df <- read.csv("fig3_E_StatTestResults_and_value.csv")

# Align order to cluster_z_score columns
mw_df <- mw_df[order(match(mw_df$Sub, names(cluster_z_score))), ]

# Define annotation color mapping
anno_colors <- c("red" = "#BF3D3D", "blue" = "#3174A1", "grey" = "grey", "NA" = "white")

# ↑↓ comparison
mw_up_down <- mw_df %>%
  mutate(sig = case_when(
    mw_up_down <= 0.05 & decreased_cluster_mean > increased_cluster_mean ~ "blue",
    mw_up_down <= 0.05 & decreased_cluster_mean < increased_cluster_mean ~ "red",
    TRUE ~ "NA"
  ))

bar_up_down <- HeatmapAnnotation(
  which = "column",
  show_legend = FALSE,
  `Increased vs Decreased` = mw_up_down$sig,
  col = list(`Increased vs Decreased` = anno_colors),
  simple_anno_size = unit(0.2, "cm"),
  name = "Increased vs Decreased",
  annotation_name_gp = gpar(fontsize = 6),
  gp = gpar(col = "black")
)

# ↑– comparison
mw_up_no <- mw_df %>%
  mutate(sig = case_when(
    mw_up_no <= 0.05 & unchanged_cluster_mean > increased_cluster_mean ~ "grey",
    mw_up_no <= 0.05 & unchanged_cluster_mean < increased_cluster_mean ~ "red",
    TRUE ~ "NA"
  ))

bar_up_no <- HeatmapAnnotation(
  which = "column",
  show_legend = FALSE,
  `Increased vs Unchanged` = mw_up_no$sig,
  col = list(`Increased vs Unchanged` = anno_colors),
  simple_anno_size = unit(0.2, "cm"),
  name = "Increased vs Unchanged",
  annotation_name_gp = gpar(fontsize = 6),
  gp = gpar(col = "black")
)

# ↓– comparison
mw_down_no <- mw_df %>%
  mutate(sig = case_when(
    mw_down_no <= 0.05 & unchanged_cluster_mean > decreased_cluster_mean ~ "grey",
    mw_down_no <= 0.05 & unchanged_cluster_mean < decreased_cluster_mean ~ "blue",
    TRUE ~ "NA"
  ))

bar_down_no <- HeatmapAnnotation(
  which = "column",
  show_legend = FALSE,
  `Decreased vs Unchanged` = mw_down_no$sig,
  col = list(`Decreased vs Unchanged` = anno_colors),
  simple_anno_size = unit(0.2, "cm"),
  name = "Decreased vs Unchanged",
  annotation_name_gp = gpar(fontsize = 6),
  gp = gpar(col = "black")
)

# Define color scale function
color_fun <- colorRamp2(break_set, colors)

# Prepare matrix for plotting
cluster_z_score_print <- cluster_z_score %>%
  select(-"group") %>%
  tibble::column_to_rownames("Sub") %>%
  t() %>%
  as.data.frame()

# Define labels
column_split <- rep(group_Set_print, times = ncol(cluster_z_score_print))
column_labels <- colnames(cluster_z_score_print)
row_labels <- rownames(cluster_z_score_print)

# Plot heatmap
library(ComplexHeatmap)
heatmap_object <- Heatmap(as.matrix(cluster_z_score_print),
  cluster_columns = TRUE,
  cluster_rows = FALSE,
  column_split = column_split,
  border_gp = gpar(col = "black", lwd = 0.3),
  rect_gp = gpar(col = "black", lwd = 0.3),
  column_labels = column_labels,
  row_labels = row_labels,
  column_names_rot = 90,
  column_names_gp = gpar(fontsize = 6),
  row_names_gp = gpar(fontsize = 6),
  clustering_distance_columns = "euclidean",
  clustering_method_columns = "ward.D2",
  row_dend_width = unit(0.5, "cm"),
  column_dend_height = unit(0.5, "cm"),
  top_annotation = c(bar_up_down, bar_up_no, bar_down_no),
  col = color_fun,
  heatmap_legend_param = list(
    title = "Correlation",
    title_position = "topcenter",
    title_gp = gpar(fontsize = 6, fontface = "plain"),
    at = break_set,
    legend_position = 'top',
    legend_direction = 'vertical',
    grid_width = unit(5, "mm"),
    grid_height = unit(10, "mm"),
    legend_height = unit(2.5, "cm"),
    legend_width = unit(0.5, "cm"),
    labels_gp = gpar(fontsize = 6)
  )
)

heatmap_object <- draw(heatmap_object)
```

## (F) CAZy gene characteristics of clusters with different abundance trends in the T2D group
## Data source: fig3_F.csv
## Plot generated using R

```{r}
# Clean workspace
rm(list = ls()) 

# Fig3_F
group_Set_print <- "T2D"
colors <- c("black", "white", "#DC8418")
break_set <- c(-2.5, 0, 2.5)

# Load data
cluster_z_score <- read.csv("fig3_F_GeneProportion.csv")
mw_df <- read.csv("fig3_F_StatTestResults_and_value.csv")

###same as above
```
## (G) KO genes related to PTS pathway in clusters with different abundance trends in the Overweight group
## Data source: fig3_G.csv
## Plot generated using R

```{r}
# Clean workspace
rm(list = ls()) 

# Fig3_G
group_Set_print <- "Overweight"
colors <- c("black", "white", "#DC8418")
break_set <- c(-2.5, 0, 2.5)

# Load data
cluster_z_score <- read.csv("fig3_G_GeneProportion.csv")
mw_df <- read.csv("fig3_G_StatTestResults_and_value.csv")

###same as above

```

## (H) KO genes related to PTS pathway in clusters with different abundance trends in the T2D group
## Data source: fig3_H.csv
## Plot generated using R

```{r}
# Clean workspace
rm(list = ls()) 

# Fig3_H
group_Set_print <- "T2D"
colors <- c("black", "white", "#DC8418")
break_set <- c(-2.5, 0, 2.5)

# Load data
cluster_z_score <- read.csv("fig3_H_GeneProportion.csv")
mw_df <- read.csv("fig3_H_StatTestResults_and_value.csv")

###same as above

```
